---
title: "Comparison of the Transcriptomic Atlas of the Human Retina and Mouse Retina"
output: html_notebook
---
# Setup
Load libraries
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(Matrix)
library(Seurat)
library(cowplot)
library(patchwork)
```
Process Human Data
```{r}
import_remote_data <- function(file_url, type = "table", header = FALSE) {
  con <- gzcon(url(file_url))
  txt <- readLines(con)
  if (type == "MM") { return (readMM(textConnection(txt))) }
  if (type == "table") { return (read.table(textConnection(txt), header = header)) }
}

count_matrix_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_counts.mtx.gz"
gene_names_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_gene_names.txt.gz"
sample_annotations_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_sample_annotations.tsv.gz"

count_matrix <- as.matrix(import_remote_data(count_matrix_URL, type = "MM"))
gene_names <- import_remote_data(gene_names_URL, type = "table")
sample_annotations <- import_remote_data(sample_annotations_URL, type = "table", header = TRUE)
```
```{r}
rownames(count_matrix) <- tolower(gene_names[,1])
colnames(count_matrix) <- tolower(sample_annotations[,1])

human_ret_seurat <- CreateSeuratObject(counts = count_matrix, 
                                       meta.data = sample_annotations, 
                                       project = "human_ret", 
                                       min.cells = 3, 
                                       min.features = 200)
human_ret_seurat
```

Process Mouse Data
```{r}
mouse.data <- Read10X(data.dir = "filtered_feature_bc_matrix")
dimnames(mouse.data)[[1]] <- tolower(dimnames(mouse.data)[[1]])
dimnames(mouse.data)[[2]] <- tolower(dimnames(mouse.data)[[2]])
mouse_ret_seurat <- CreateSeuratObject(counts = mouse.data, 
                                       project = "mouse_ret", 
                                       min.cells = 3, 
                                       min.features = 200)
mouse_ret_seurat
```
Combine
```{r}
ret.list <- list(human = human_ret_seurat, mouse = mouse_ret_seurat)

ret.list <- lapply(X = ret.list, FUN = function(x) {
    x <- NormalizeData(x, verbose = FALSE)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
})
```

# Integration
```{r}
ret.anchors <- FindIntegrationAnchors(object.list = ret.list, dims = 1:50,  anchor.features = 1000)
ret.combined <- IntegrateData(anchorset = ret.anchors, dims = 1:50)
```

# Integrated Analysis
```{r}
DefaultAssay(ret.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
ret.combined <- ScaleData(ret.combined, verbose = FALSE)
ret.combined <- RunPCA(ret.combined, npcs = 50, verbose = FALSE)
# t-SNE and Clustering
ret.combined <- RunUMAP(ret.combined, reduction = "pca", dims = 1:35)
ret.combined <- FindNeighbors(ret.combined, reduction = "pca", dims = 1:35)
ret.combined <- FindClusters(ret.combined, resolution = 0.07)
```
# UMAP Visualization
```{r}
# Visualization
DimPlot(ret.combined, reduction = "umap", group.by = "orig.ident")
DimPlot(ret.combined, reduction = "umap", label = TRUE)
```
```{r}
DimPlot(ret.combined, reduction = "umap", split.by = "orig.ident")
```

# Identify Conserved Markers
```{r}
max_cluster <- 8
marker.list <- as.list(c(0:max_cluster))
marker.list <- lapply(X = marker.list, FUN = function(x) {
  x <- FindConservedMarkers(ret.combined, ident.1 = x, grouping.var = "orig.ident", verbose = TRUE)
})
```
```{r}
DefaultAssay(ret.combined) <- "RNA"

features <- tolower(c("Pde6a","Gnat2","Nefl","Camk2b","Thy1","Gad1","Slc6a9","Pcsk6","Trpm1","Sept4","Glul","Arr3","C1qa","Tm4sf1"))

p <- FeaturePlot(object = ret.combined, 
            features = features, 
            pt.size = 0.07,
            cols = c("lightgrey", "#F26969"),
            min.cutoff = "q9",
            combine = FALSE)

for(i in 1:length(p)) {
  p[[i]] <- p[[i]] + NoLegend() + NoAxes()
}

cowplot::plot_grid(plotlist = p, ncol=3)
```

* Rod : pde6a
* AC (amacrine cell) : gad1, slc6a9
* MG (Müller glia) : glul
* BC (bipolar cell) : Trpm
* CC (cone cell) : gnat2, arr3
* RGC (retinal ganglial cell) : nefl, thy1
* VC (vascular cell) : mgp, tm4sf1
* M (microglia) : c1qa
* HC (horizontal cell) : sept4

Markers were determined from [this](https://www.nature.com/articles/s41467-019-12780-8) paper and other sources.
```{r}
ret.combined <- RenameIdents(ret.combined, `0` = "Rod", `1` = "BC", `2` = "MG", 
    `3` = "MG", `4` = "RGC", `5` = "CC", `6` = "AC", `7` = "VC", `8` = "HC", `9` = "M", 
    `10` = "HC")

DimPlot(ret.combined, label = TRUE)
```




