---
title: "Integrating Primate Data into Analysis"
output: html_notebook
---
# Setup
Load libraries
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(Matrix)
library(Seurat)
library(cowplot)
library(patchwork)

# parallelization
library(future)
options(future.globals.maxSize= +Inf)
plan()
```
Process Human Data
```{r}
import_remote_data <- function(file_url, type = "table", header = FALSE) {
  con <- gzcon(url(file_url))
  txt <- readLines(con)
  if (type == "MM") { return (readMM(textConnection(txt))) }
  if (type == "table") { return (read.table(textConnection(txt), header = header)) }
}
count_matrix_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_counts.mtx.gz"
gene_names_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_gene_names.txt.gz"
sample_annotations_URL <- "https://ftp.ncbi.nlm.nih.gov/geo/series/GSE137nnn/GSE137537/suppl/GSE137537_sample_annotations.tsv.gz"

human.count_matrix <- as.matrix(import_remote_data(count_matrix_URL, type = "MM"))
human.gene_names <- import_remote_data(gene_names_URL, type = "table")
human.sample_annotations <- import_remote_data(sample_annotations_URL, type = "table", header = TRUE)
```
```{r}
rownames(human.count_matrix) <- tolower(human.gene_names[,1])
colnames(human.count_matrix) <- tolower(human.sample_annotations[,1])

human_ret_seurat <- CreateSeuratObject(counts = human.count_matrix, 
                                       meta.data = human.sample_annotations, 
                                       project = "human_ret", 
                                       min.cells = 3, 
                                       min.features = 200)
human_ret_seurat
```

Process Mouse Data
```{r}
mouse.data <- Read10X(data.dir = "filtered_feature_bc_matrix")
dimnames(mouse.data)[[1]] <- tolower(dimnames(mouse.data)[[1]])
dimnames(mouse.data)[[2]] <- tolower(dimnames(mouse.data)[[2]])
mouse_ret_seurat <- CreateSeuratObject(counts = mouse.data, 
                                       project = "mouse_ret", 
                                       min.cells = 3, 
                                       min.features = 200)
mouse_ret_seurat
```

Process Primate Data
```{bash}
url=https://ftp.ncbi.nlm.nih.gov/geo/series/GSE118nnn/GSE118546/suppl/GSE118546_macaque_fovea_all_10X_Jan2018.Rdata.gz
wget $url -O primate_data/GSE118546_macaque_fovea_all_10X_Jan2018.Rdata.gz
gunzip primate_data/*
```
```{r}
install.packages( c('devtools', 'roxygen2') )
library(devtools)
library(roxygen2)
install_github( 'hb-gitified/cellrangerRkit',
                auth_token = 'your_token' )
```
```{r}
load("primate_data/GSE118546_macaque_fovea_all_10X_Jan2018.Rdata")

dimnames(Count.mat_fovea)[[1]] <- tolower(dimnames(Count.mat_fovea)[[1]])
macaque_fovea_seurat <- CreateSeuratObject(Count.mat_fovea,
                                           project = "macaque_fovea", 
                                           min.cells = 3, 
                                           min.features = 200)

# give macaque dta uniform name in "orig.ident" metadata column
AddMetaData(macaque_fovea_seurat, 
            metadata = macaque_fovea_seurat[["orig.ident"]], 
            col.name = "orig.sample.name")
macaque_fovea_seurat[["orig.ident"]] <- "macaque_fovea"

macaque_fovea_seurat
```
Cleanup
```{r}
rm(human.count_matrix, human.gene_names, human.sample_annotations)
rm(count_matrix_URL, gene_names_URL, sample_annotations_URL, import_remote_data)
rm(mouse.data)
rm(Count.mat_fovea, macaque_fovea)
```


Combine
```{r}
# combine
ret.list <- list(human = human_ret_seurat, mouse = mouse_ret_seurat, macaque = macaque_fovea_seurat)

# preprocess
ret.list <- lapply(X = ret.list, FUN = function(x) {
    x <- NormalizeData(x, verbose = FALSE)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
})

# cleanup
rm(human_ret_seurat, mouse_ret_seurat, macaque_fovea_seurat)
```

# Integration
```{r}
plan("multiprocess", workers = 4)
ret.anchors <- FindIntegrationAnchors(object.list = ret.list, dims = 1:50,  anchor.features = 1000)
plan("multiprocess", workers = 1)
ret.combined <- IntegrateData(anchorset = ret.anchors, dims = 1:50)
```

# Integrated Analysis
```{r}
plan("multiprocess", workers = 4)

DefaultAssay(ret.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
ret.combined <- ScaleData(ret.combined, verbose = FALSE)
ret.combined <- RunPCA(ret.combined, npcs = 50, verbose = FALSE)
# t-SNE and Clustering
ret.combined <- RunUMAP(ret.combined, reduction = "pca", dims = 1:35)
ret.combined <- FindNeighbors(ret.combined, reduction = "pca", dims = 1:35)
ret.combined <- FindClusters(ret.combined, resolution = 0.075)
```
# UMAP Visualization
```{r warning=FALSE}
DimPlot(ret.combined, reduction = "umap", group.by = "orig.ident")
DimPlot(ret.combined, reduction = "umap", label = TRUE)
```
```{r, fig.height = 4, fig.width = 3}
DimPlot(ret.combined, reduction = "umap", split.by = "orig.ident", ncol = 1)
```

# Identify Clusters with Canonical Markers
```{r}
DefaultAssay(ret.combined) <- "RNA"

features <- tolower(c("Pde6a","Gnat2","Nefl","Camk2b","Thy1","Gad1","Slc6a9",
                      "Pcsk6","Trpm1","Sept4","Glul","Arr3","C1qa","Tm4sf1", "Mgp"))

FeaturePlot(object = ret.combined, 
            features = features, 
            pt.size = 0.1,
            cols = c("lightgrey", "#F26969"),
            min.cutoff = "q9",
            combine = TRUE) & NoLegend() & NoAxes

# Cowplot method: make sure to change to "combine = FALSE" and remove "& NoLegend() & NoAxes"

# for(i in 1:length(p)) {
#   p[[i]] <- p[[i]] + NoLegend() + NoAxes()
# }
# 
# cowplot::plot_grid(plotlist = p, ncol=3)
```

* Rod : pde6a
* AC (amacrine cell) : gad1, slc6a9
* MG (Müller glia) : glul
* BC (bipolar cell) : Trpm, camk2b
* CC (cone cell) : gnat2, arr3
* RGC (retinal ganglial cell) : nefl, thy1
* VC (vascular cell) : mgp, tm4sf1
* M (microglia) : c1qa
* HC (horizontal cell) : sept4

Markers were determined from [this](https://www.nature.com/articles/s41467-019-12780-8) paper and other sources.
```{r}
ret.combined <- RenameIdents(ret.combined, `0` = "MG", `1` = "Rod", `2` = "RGC", 
    `3` = "RGC", `4` = "BC", `5` = "CC", `6` = "BC", `7` = "AC", `8` = "BC", `9` = "RGC", 
    `10` = "RGC", `11`= "HC", `12` = "MG", `13` = "VC", `14` = "RGC", `15` = "RGC", `16` = "M", `17` = "RGC")

DimPlot(ret.combined, label = TRUE)
```


# Find Differentially Expressed Genes
```{r}
cells.types <- c("Rod", "BC", "MG", "RGC", "CC", "AC", "VC", "HC", "M")
theme_set(theme_cowplot())

cell_type_avg <- function(seurat.combined, ident) {
  cells.x <- subset(seurat.combined, idents = ident)
  Idents(cells.x) <- "orig.ident"
  cells.x.avg <- log1p(AverageExpression(cells.x, verbose = FALSE)$RNA)
  cells.x.avg$gene <- rownames(cells.x.avg)
  return(cells.x.avg)
}

cells.plot <- as.list(cells.types)
cells.plot <- lapply(cells.plot, FUN = function(x) {
  cells.x.avg <- cell_type_avg(ret.combined, ident = x)
  x <- ggplot(cells.x.avg, aes(human_ret, mouse_ret)) + geom_point(size = 0.1) + ggtitle(x)
  return(x)
})

# For individual plots
# for (p in cells.plot) {
#   print(p)
# }

# For grid plot
cowplot::plot_grid(plotlist = cells.plot, ncol = 3)
```
```{r}
ret.combined$celltype.organism <- paste(Idents(ret.combined), ret.combined$orig.ident, sep = "_")
ret.combined$celltype <- Idents(ret.combined)
Idents(ret.combined) <- "celltype.organism"
```
```{r}
cells.diffgenes <- as.list(cells.types)
cells.diffgenes <- lapply(cells.diffgenes, FUN = function(x) {
  lab_human <- sprintf("%s_human_ret", x)
  lab_mouse <- sprintf("%s_mouse_ret", x)
  return(FindMarkers(ret.combined, ident.1 = lab_human, ident.2 = lab_mouse, verbose = FALSE))
})
```
Tables with the most differentially expressed genes in each cell subtype:
```{r}
for(i in seq_along(cells.diffgenes)) {
  print(knitr::kable(head(cells.diffgenes[[i]]),caption=cells.types[[i]]))
}
```
Save as csv files
```{r}
for(i in seq_along(cells.diffgenes)) {
  write.csv(cells.diffgenes[[i]], sprintf("results/%d_%s.csv", i, cells.types[[i]]))
}
```

```{r warning=FALSE}
genes_to_plot <- 3
for (i in seq_along(cells.types)) {
  print(FeaturePlot(object = ret.combined, 
              features = rownames(cells.diffgenes[[i]])[1:genes_to_plot], 
              split.by = "orig.ident", 
              max.cutoff = 3, 
              cols = c("grey", "red"),
              pt.size = 0.07,
              combine = TRUE,
              label.size = 0.5
              ) + plot_annotation(title = cells.types[[i]]) & NoLegend() & NoAxes()
        )
}
```

Check cell proportion for each species:
```{r}
knitr::kable(prop.table(x = table(Idents(ret.combined), ret.combined@meta.data$orig.ident), margin = 2))
```

